Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology

### 2024-11-06 analysis/paper/article.qmd
minor edits

Our results reveal some of these shifts, providing insights into the dynamics of topical, methods, and geographic foci over time. While these dynamics are deeply felt by people actively seeking faculty jobs, we believe this is the first time they have been quantified at such a large scale. Trends in job ads reflect broader shifts in intellectual and practical priorities about archaeology, undergraduate education, and the process of hiring professors. The demand for archaeology faculty, indicated by the total number of tenure-track jobs, maybe explained by a variety of processes. Overall we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-14. This is at odds with prior work finding that contingent positions make up more than half of faculty positions. This discrepancy may be due to advertising for short-term positions typically having much more limited circulation than advertising for tenure-track jobs, and thus less likely to be added to the Academic Jobs Wiki. The downward trend in tenure-track positions during 2013-2019 maybe related to decline in undergraduate enrollment in anthropology fields since 2013 [@crambChangingProfileTenureTrack2022a]. The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities focused on adapting to online instruction and assessment in an effort to minimize the spread of the virus.

@@

A key limitation of our research is that it does not include an analysis of the academic profiles of those who were eventually hired from the job ads. Our results reveal collective aspirations for the future of the field, but only from those who are securely employed in it as tenured professors serving on hiring committees at US universities. Our results do not show how these aspirations worked out in the topical and geographic foci of the people who were eventually hired for the positions studied here. Future work should consider interviewing faculty hired during our study period to match up people to the ads they responded to. If the successful applicants can be identified, then the match between the details of the job ad and the applicant's research can be analysed. Then we can assess how effective the hiring process is for driving change in the discipline by setting the topics of courses that will be taught in undergraduate and graduate curricula, and research that will be supported by universities.

>>>>

Our results reveal some of these shifts, providing insights into the dynamics of topical, methods, and geographic foci over time. While these dynamics are deeply felt by people actively seeking faculty jobs, we believe this is the first time they have been quantified at such a large scale. Trends in job ads reflect broader shifts in intellectual and practical priorities about archaeology, undergraduate education, and the process of hiring professors. The demand for archaeology faculty, indicated by the total number of tenure-track jobs, maybe explained by a variety of processes. Overall we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-14. This is at odds with prior work finding that contingent positions make up more than half of faculty positions. This discrepancy may be due to advertising for short-term positions typically having much more limited circulation than advertising for tenure-track jobs, and thus less likely to be added to the Academic Jobs Wiki. Many of these short-term positions are not advertised nationally, but only through local email lists and are filled by people close to the hiring department, such as recently graduated students. The downward trend in tenure-track positions during 2013-2019 maybe related to decline in undergraduate enrollment in anthropology fields since 2013 [@crambChangingProfileTenureTrack2022a]. The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities focused on adapting to online instruction and assessment in an effort to minimize the spread of the virus.

@@

A key limitation of our research is that it does not include an analysis of the academic profiles of those who were eventually hired from the job ads. Our results reveal collective aspirations for the future of the field, but only from those who are securely employed in it as tenured professors serving on hiring committees at US universities. Our results do not show how these aspirations worked out in the topical and geographic foci of the people who were eventually hired for the positions studied here. Future work should consider interviewing faculty hired during our study period to match up people to the ads they responded to. If the successful applicants can be identified, then the match between the details of the job ad and the applicant's research can be analysed. Then we can assess how effective the hiring process is for driving change in the discipline by setting the topics of courses that will be taught in undergraduate and graduate curricula, and research that will be supported by universities. We also recognize the possibility that the Academic Jobs Wiki might not capture all available positions or might over-represent certain types of institutions.
